Monte Carlo and bootstrap for ALL parameters with tables

### 2024-09-04 All3_tables_grid_copy.Rmd
With or without stratified boot

@@

@@

MC_boot <- function(n, B1, B2, f_sampling){
  # n: sample size;
  # B1: number of iterations for MC;
  # B2: number of iterations for bootstrap

  p <- sum(rep((1/1000)*(1/1000), n))
  BOOT_df <- list() # All local variables
  BOOT_est <- list() # All estimations
  BOOT_var <- list() # Variances
  BOOT_CI <- list() # Confidence Intervals
  for (b in 1:B1) { # One MC iteration at a time
    # Sampling in the territory
    df_grid <- f_sampling(i,c_dim)
    xy <- data.frame(a1 = df_grid$x, a2 = df_grid$y)
    # Variables of interest derived from the sample
    all_var <- as.data.frame(t(mapply(circle3_ALL_, xy$a1, xy$a2)))
    nb_var <- length(all_var) # number of variables of interest
    df_b <- cbind(xy, all_var)

@@

    # Bootstrap iterations
    df_boot <- f_boot(df_b, B2, nb_var)
    est_boot_val <- est_boot(df_boot, p, nb_var)

@@

# Example: Algorithm 8

@@

>>>>

```{r}
## Making of bootstrap samples based on an initial sample
f_boot_strat <- function(df_sample, B, nb_var){
  # df_sample: initial sample with nb_est variables and of size n; 
  # B: number of bootstrap samples made thanks to df_sample;
  # nb_var: number of variables of interest (for totals and not ratios)

  df_boot <- data.frame(useless = rep(0, nrow(df_sample)/2)) # Only one out of 2 values
  for (b in 1:B) {
    nb <- c()
    for (stratum in 1:(nrow(df_sample)/2)) {
      strat_res <- sample(c(stratum*2-1, stratum*2), 1)
      nb <- append(nb, strat_res)
    }
    # (n/2) points at random
    sample_b <- df_sample[nb,]
    df_boot <- cbind(df_boot, sample_b)
    colnames(df_boot)[((b-1)*(nb_var+2)+1 + 1):(b*(nb_var+2)+1)] <- as.vector(mapply(function(j){paste("ech", b, colnames(df_sample)[j], sep = "")}, 1:(nb_var+2)))
  }
  df_boot$useless <- NULL
  return(df_boot)
  # Return a data frame with B samples of (nb_var+2) variables and of size (n/2)
}
```

@@

```{r}
## Making of estimations for each sample made by the bootstrap method
est_boot_strat <- function(df, p, nb_var){
  # df: data frame with (nb_var+2)*B columns and (n-1) rows
  # p: inclusion density function (constant in the uniform case)
  # nb_var: number of variables of interest (for totals and not ratios)

  n <- nrow(df)*2 # sample size of the initial sample
  nboot <- length(df)/(nb_var+2) # number of bootstrap samples
  big_est <- as.data.frame(matrix(NA, ncol = nb_var, nrow = nboot))

  # Estimations of totals
  for (c in 1:nboot) {
    big_est <- mapply(function(vec_, j){
        vec_[c] <- (2)*sum(df[, 2 + j + ((nb_var+2)*(c-1))]/p)
        return(as.data.frame(vec_))
        }, big_est, 1:nb_var)
    # (2) is the correction needed since we choose only (n/2) points
  }
  # Estimations of ratios
  l_big_est <- length(big_est)
  big_est <- as.data.frame(big_est)
  for (k in 1:4) {
    big_est <- add_column(big_est, rep(NA, nrow(big_est)))
  }
  for (c in 1:nboot) {
    ratio_c <- mapply(function(j){big_est[c,j+4]/big_est[c,j]
    }, 1:4) # 4 ratios
    big_est[c, (l_big_est+1):(l_big_est+4) ] <- ratio_c
  }
  colnames(big_est) <- paste0("est_", raw_names)[-(1:2)]
  return(big_est)
  # Data frame with estimations for each sample
}
```

@@

MC_boot <- function(n, B1, B2, f_sampling, strat=FALSE){ # strat=c('TRUE','FALSE')
  # n: sample size;
  # B1: number of iterations for MC;
  # B2: number of iterations for bootstrap
  # strat: which type of bootstrap

  if (strat == FALSE){
    fun_f_boot <- f_boot
    fun_est_boot <- est_boot
  }else{
    fun_f_boot <- f_boot_strat
    fun_est_boot <- est_boot_strat
  }
  p <- sum(rep((1/1000)*(1/1000), n))
  BOOT_df <- list() # All local variables
  BOOT_est <- list() # All estimations
  BOOT_var <- list() # Variances
  BOOT_CI <- list() # Confidence Intervals
  for (b in 1:B1) { # One MC iteration at a time
    # Sampling in the territory
    df_grid <- f_sampling(i,c_dim)
    xy <- data.frame(a1 = df_grid$x, a2 = df_grid$y)
    # Variables of interest derived from the sample
    all_var <- as.data.frame(t(mapply(circle3_ALL_, xy$a1, xy$a2)))
    nb_var <- length(all_var) # number of variables of interest
    df_b <- cbind(xy, all_var)

@@

    # Bootstrap iterations
    df_boot <- fun_f_boot(df_b, B2, nb_var)
    est_boot_val <- fun_est_boot(df_boot, p, nb_var)

@@

# Example: Algorithm 8 with basic bootstrap

@@

# Example: Algorithm 8 with stratified bootstrap

res2 <- MC_boot(n, B1, B2, fun_samp, strat = TRUE)
# Just one circle to begin with

```{r RS_algo8_strat}
tab_rs <- DF_rs %>%
  kable("latex", booktabs = FALSE, align = "lc", linesep = c("", "", "",
  "\\hline"), caption = paste0("Relative stability of variance by using ", B1, " Monte Carlo iterations and ", B2, " bootstrap replications"), col.names = gsub("[.]", " ", names(DF_rs))) %>% # \\hline \\addlinespace
  kable_styling(latex_options = c("striped", "hold_position")) %>% # bootstrap_options = "striped", full_width = FALSE, 
  row_spec(0, bold = TRUE) %>% # , hline_after = TRUE
  row_spec(c(0, 4, 8), hline_after = T) %>% 
  collapse_rows(columns = 1, latex_hline = "major")
```

```{r CI_tot_algo8_strat}
tab_ci1 <- df_total1 %>%
  kable("latex", booktabs = FALSE, align = "llcccc", caption = paste0("Characteristics of confidence intervals when using ", B1, " Monte Carlo iterations and ", B2, " bootstrap replications (totals)")) %>%
  kable_styling(latex_options = "hold_position") %>%
  row_spec(0, bold = TRUE) %>%
  row_spec(c(0, 16), hline_after = T) %>% 
  row_spec(c(16), hline_after = T) %>% 
  collapse_rows(columns = 1, valign = "middle")
```

```{r CI_ratio_algo8_strat}
tab_ci2 <- df_total2 %>%
  kable("latex", booktabs = FALSE, align = "llcccc", caption = paste0("Characteristics of confidence intervals when using ", B1, " Monte Carlo iterations and ", B2, " bootstrap replications (ratios)")) %>%
  kable_styling(latex_options = "hold_position") %>%
  row_spec(0, bold = TRUE) %>%
  row_spec(c(0), hline_after = T) %>% 
  # row_spec(c(16,32), hline_after = T) %>% 
  collapse_rows(columns = 1, valign = "middle")
```

```{r RB_algo8_strat}
tab_var <- diff_var %>%
  kable("latex", booktabs = FALSE, align = "lc", linesep = c("", "", "",
  "\\addlinespace"), caption = paste0("Approximation of Relative Bias when using ", B1, " Monte Carlo iterations and ", B2, " bootstrap replications in one case and ", B0, " Monte Carlo iterations in the other case"), col.names = gsub("[.]", " ", names(diff_var))) %>% # \\hline
  kable_styling(latex_options = c("striped", "hold_position")) %>% # bootstrap_options = "striped", 
  row_spec(0, bold = TRUE) %>% # , hline_after = TRUE
  row_spec(c(0, 4, 8), hline_after = T) %>% 
  collapse_rows(columns = 1, latex_hline = "major")
```
